Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-05-14 SupportingInformation.Rmd
some changes

<!--Another model, distinct from the above, Dan and Shya to discuss-->
```{r asym_sens_beta_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
source("asym_sensitivity.R")

#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       N : number of global and local noise points you want
#       rho : covariance of global noise matrix
#       alpha : a number : local noise coeeficient
#       beta : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

# Output :
#     a list of 4 pop matrices 
#                         1. (for beta=+ve, decrease=F)
#                         2. (for beta=+ve, decrease=T)
#                         3. (for beta=-ve, decrease=F)
#                         4. (for beta=-ve, decrease=T)

plotter_asym_sens_beta_varn<-function(N,rho,alpha,beta,numkeep_last,ploton){

  pop<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  neg_beta<- -beta
  pop_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

@@

    plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    mtext(paste0("(A)"),side = 3, line=0.3, adj=0.3, col="black")
    plot(pop_d[,1],pop_d[,2],col="blue",main = bquote(beta ==.(beta)))
    mtext(paste0("(B)"),side = 3, line=0.3, adj=0.3, col="black")
    plot(pop_neg_beta[,1],pop_neg_beta[,2],col="orange",main = bquote(beta ==.(neg_beta)))
    mtext(paste0("(C)"),side = 3, line=0.3, adj=0.3, col="black")
    plot(pop_d_neg_beta[,1],pop_d_neg_beta[,2],col="skyblue",main = bquote(beta ==.(neg_beta)))
    mtext(paste0("(D)"),side = 3, line=0.3, adj=0.3, col="black")

@@

  return(list(pop=pop,
              pop_d=pop_d,
              pop_neg_beta=pop_neg_beta,
              pop_d_neg_beta=pop_d_neg_beta))

}
#--------------------------------------------------------------
N<-10000
rho<-0.8
alpha<-0.2
numkeep_last<-1000

pdf("./Results/asym_sens_results/beta_0.1_asym_sens.pdf")
ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
dev.off()

pdf("./Results/asym_sens_results/beta_0.9_asym_sens.pdf")
ans_beta_0.9<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
dev.off()

vlist_beta_0.1_beta_0.9<-c(ans_beta_0.1,ans_beta_0.9) # a list of 8 matrices

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AIC_w","relLTdep_AICw","relUTdep_AIC_w","p_CvM","p_KS","Numboot_success")
bg <- intToUtf8(0x03B2) 
cnm<-c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T")
cname<-paste0(bg,cnm)
colnames(summary_stat)<-cname

stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
names(stat_npa_asym_beta_0.1_beta_0.9)<-cname

for(ic in c(1:8)){
  v<-vlist_beta_0.1_beta_0.9[[ic]]
  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                      gofnormal=F,status=F)

@@

  stat_npa_asym_beta_0.1_beta_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cname[ic],numbin=10)

@@

saveRDS(summary_stat,"./Results/asym_sens_results/beta_variation_summary_stat.RDS")
saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
```

![results for model selection \label{fig_ms_asym_beta_0.1}](./Results/asym_sens_results/beta_0.1_asym_sens.pdf){ width=40%, height=40% }

![results for model selection \label{fig_ms_asym_beta_0.9}](./Results/asym_sens_results/beta_0.9_asym_sens.pdf){ width=40%, height=40% }

```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
cnm2<-c("b_0.1","b_0.1d_T","b_-0.1","b_-0.1d_T","b_0.9","b_0.9d_T","b_-0.9","b_-0.9d_T")
colnames(summary_stat)<-cnm2
knitr::kable(summary_stat, booktabs = T, 
             format = "pandoc",
             digits=3,
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption="summary of model selection results for asym. sensitivity \\label{tab_summary_stat_asym_sens}")
```

\newpage
```{r tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")

for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  temp2<-knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste("beta",cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
  print(temp2)
  cat("\\newpage")
}
```

>>>>

<!--Another model, distinct from the above, Dan and Shya to discuss-->
```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
#-----------------------------------------------------------------------------------
source("OurBiCopSelect.R")
source("bivfunctionplot.R")
source("asym_sensitivity.R")
#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       mg : global noise matrix (dim = N by 2)
#       ml : local noise matrix (dim = N by 2)
#       a : a number : local noise coeeficient
#       b : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

# Output :
#     a list of 4 pop matrices 
#                         1. (for b=+ve, decrease=F)
#                         2. (for b=+ve, decrease=T)
#                         3. (for b=-ve, decrease=F)
#                         4. (for b=-ve, decrease=T)

plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton){

  pop<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  neg_b<- -b
  pop_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

@@

    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    mtext(expression(paste("(A) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    mtext(expression(paste("(B) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")
    plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    mtext(expression(paste("(C) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
    plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    mtext(expression(paste("(D) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")

@@

  return(list(pop=pop,
              pop_d=pop_d,
              pop_neg_b=pop_neg_b,
              pop_d_neg_b=pop_d_neg_b))

}
#----------------------------------------------------------
N<-10000
r<-0.8
a<-0.2
numkeep_last<-1000
temp<-get_noise_glob_loc_mat(N=N,r=r)
mg<-temp$mg
ml<-temp$ml

pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf")
ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf")
ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T)
dev.off()

vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_0.9) # a list of 8 matrices

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","Numboot_success")
#bg <- intToUtf8(0x03B2) 
cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
#cname<-paste0(bg,cnm)
colnames(summary_stat)<-cnm

ms_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
names(ms_asym_b_0.1_b_0.9)<-cnm

stat_npa_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
names(stat_npa_asym_b_0.1_b_0.9)<-cnm

for(ic in c(1:8)){
  v<-vlist_b_0.1_b_0.9[[ic]]
  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                      gofnormal=F,status=F)

  ms_asym_b_0.1_b_0.9[[ic]]<-res

@@

  stat_npa_asym_b_0.1_b_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)

@@

saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
saveRDS(ms_asym_b_0.1_b_0.9,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.9.RDS")
saveRDS(stat_npa_asym_b_0.1_b_0.9,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
```

![results for model selection \label{fig_ms_asym_b_0.1}](./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf){ width=40%, height=40% }

![results for model selection \label{fig_ms_asym_b_0.9}](./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf){ width=40%, height=40% }

```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
knitr::kable(summary_stat, booktabs = T, 
             format = "pandoc",
             digits=3,
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption="summary of model selection results for asym. sensitivity \\label{tab_summary_stat_asym_sens}")
```

\newpage
```{r tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")

for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  temp2<-knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
  print(temp2)
  cat("\\newpage")
}
```
